
---
title: "GIS"
author: "Shani Fourie"
date: "22/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

The project goal is to overlay marine protected areas of SA and the occurrence of the endangered African penguin to observe the density of the penguins within and near the marine protected areas.

```{r}
library(sf)
methods(class="sf")
marine=st_read("SAMPAZ_OR_2021_Q4.shp")

```


Plotting the marine protected areas of SA. 
```{r}
#Getting the coordinates
st_crs(marine)


head(marine)

#plotting
plot(marine)
plot(marine[2])
library(tidyverse)
ggplot() +
  geom_sf(data=marine, aes(fill = `CUR_NME`))

#Cropping
ext=c(14.1,-40,25,-32)
names(ext)=c("xmin","ymin","xmax","ymax")
ext


#feeding cropped area into st_crop

marine=st_crop(marine,ext)
ggplot() + geom_sf(data=marine, aes(fill = `CUR_NME`))+theme(legend.position = "none")


```






The occurrence of the African penguins around the coast of the western Cape
```{r}
#calling iNaturalist locality point data from R
library(rinat)
library(tidyverse)
pc <- get_inat_obs(taxon_name = "Spheniscus demersus",
                   bounds = c(-40,14.1,-32,25),
                   maxresults = 1000)

#view the first few rows
head(pc)
class(pc)


#Make the dataframe a spatial object of class = "sf"
pc <- st_as_sf(pc, coords = c("longitude", "latitude"), crs = 4326)

#What class is it?
class(pc)

names(pc)


#plot
ggplot() + geom_sf(data=pc)


#adding in Basemaps
library(rosm)
library(ggspatial)

ggplot() + 
  annotation_map_tile(type = "osm", progress = "none") + 
  geom_sf(data=pc)
  
```





In this leaflet you can see the observed African penguins (dots) appear to be more abundant within and near the marine protected areas (polygons)
```{r}
#Interactive maps with leaflet
library(leaflet)
library(htmltools)
library(tidyverse)

leaflet() %>%
  # Add default OpenStreetMap map tiles
  addTiles(group = "Default") %>%  
  # Add our points
  addCircleMarkers(data = pc,
                   group = "Spheniscus demersus",
                   radius = 3, 
                   color = "black") %>%
  addPolygons(data=marine, weight = 2, col=rainbow(10, alpha=.5))

```

```{r}
library(terra)
```



This plot provides the distribution of the African penguins along with the marine protected areas of South Africa
```{r}
#plot
ggplot() +  annotation_map_tile(type = "osm", progress = "none")+
  geom_sf(data=pc)+ geom_sf(data=marine, aes(fill = `CUR_NME`))+theme(legend.position = "none")


```

